---
title: "Daily streamflow and brown trout population processes in regulated Montana rivers"
---

```{r setup, include=FALSE}
source("_common.R")
```

## What this is

Analysis supporting Bureau of Reclamation flow-management decisions on the Madison
River and other regulated Montana streams. It estimates, for **every day of the
calendar year**, how streamflow affects brown trout recruitment, adult survival, and
total population production.

Three questions:

1. **When does releasing water do the most good?** Guidance for water managers,
   irrigators, and holders of leased water rights.
2. **What do daily-scale results say about trout population ecology in Montana?**
   Which vital rate is flow-limited, and in which season.
3. **How do the answers change under Bureau of Reclamation forecasted flows?**

## Pages

- **[Methods](methods.html)** — how the analysis works, and what it can and cannot resolve
- **[Results](results.html)** — one page per site: daily guidance, scenarios, and the equilibrium rule-curve arm
- **[Among sites](results_among_sites.html)** — what the daily results say about Montana brown trout ecology

## Headline

```{r headline, results='asis'}
if (is.null(GATE)) {
  cat("> **Analysis not yet run.** Execute `RUN_ALL.R` and re-render.\n")
} else {
  ok_b <- GATE_OK("recruitment", "beta"); ok_s <- GATE_OK("recruitment", "seasonal")

  # gate.csv labels the functional arms "beta:<estimator>", never plain "beta".
  # An exact match on "beta" returns nothing and renders as [PENDING].
  gr <- function(col, frm) {
    v <- GATE[[col]][GATE$process == "recruitment" & GATE$form == frm]
    if (!length(v)) NULL else v[1]
  }
  bform <- paste0("beta:", CFG$fitting$estimator)
  # V() already returns "**[PENDING]**"; do not wrap it in ** again or the
  # markdown renders four literal asterisks.
  b <- function(x) if (grepl("^\\*\\*", x)) x else paste0("**", x, "**")

  cat(sprintf("Blocked cross-validated $R^2$ for the daily curve (%s) was %s (permutation $p$ = %s); the four-window confirmatory model gave %s.\n\n",
    CFG$fitting$estimator,
    b(V(gr("r2_blocked", bform))), b(V(gr("p_value", bform))),
    b(V(gr("r2_blocked", "seasonal")))))

  # If both estimators were gated, agreement between them is the reporting
  # standard for any daily claim -- see docs/notes/FPC_reviewer_critiques.md.
  if (isTRUE(CFG$fitting$compare_estimators)) {
    other <- setdiff(c("penalized", "fpc"), CFG$fitting$estimator)
    o <- gr("r2_blocked", paste0("beta:", other))
    if (!is.null(o))
      cat(sprintf("The %s estimator gave %s on the same folds. %s\n\n", other, b(V(o)),
        if (isTRUE(ok_b)) "Both arms are reported; agreement between them is the standard for a daily claim."
        else "Neither arm clears the gate."))
  }
  if (isTRUE(ok_b)) {
    cat("The daily curve predicts out of sample, so timing guidance is reported below.\n")
  } else if (isTRUE(ok_s)) {
    cat("**Flow affects recruitment, but the daily timing is not resolvable in this record.** Guidance is reported at seasonal resolution.\n")
  } else {
    cat("**No flow signal survives validation.** Results are reported as a null. Curves are shown for completeness and labelled unsupported throughout.\n")
  }
}
```

```{r iv, results='asis'}
if (!is.null(IVS)) {
  cat("\n### Resolvable release windows\n\n")
  IVS %>%
    transmute(Process = process, Which = which,
              Window = paste0(doy_date(lo_doy), " – ", doy_date(hi_doy)),
              `Width (days)` = round(width_days),
              Resolvable = ifelse(resolvable, "yes", "no — report a season"),
              `Gate passed` = ifelse(passes, "yes", "NO")) %>%
    kable() %>% print()
}
```

---

*Reproducible: every number and figure on this site is generated by the scripts in
`R/` and read from `output/`. Source at the repository linked in the navigation
bar.*
