---
title: "Materials and Methods"
subtitle: "Daily streamflow and brown trout population processes"
---

```{r setup, include=FALSE}
source("_common.R")
```

<!--
PNAS Research Articles place Materials and Methods after the Results and cap the
main text near 4,000 words including everything, so a Methods section much past
~1,300 words will not fit. Detail that does not fit belongs in SI Appendix, which
is the second half of this document.

Sections marked [VERIFY] contain an assumption that could not be checked from the
code. Each is listed in PROJECT_LOG.md as an open item.
-->

# Materials and Methods

## Study system

Brown trout (*Salmo trutta*) populations in regulated Montana rivers, with the
Madison River below Ennis Lake (USGS gage 06041000) as the focal site. Discharge is
managed under Bureau of Reclamation operations. Analysis covers
`r Vi(S[[paste0("n_years_", CFG$sites$active[[1]])]])` recruit-years.

*[VERIFY: state the gage-to-reach distance, any intervening accretion, and the
long-term monitoring design — reach lengths, sampling season, years — with a
citation to the Montana Fish, Wildlife and Parks program.]*

## Population model and the response variables

Abundance and demographic rates were estimated with a Bayesian age-structured
integrated population model fitted in JAGS. Mark–recapture abundance estimation is
conducted separately, under a protocol agreed with Montana Fish, Wildlife and
Parks so that estimates match published values. The model tracks age-2 recruits
($N_2$), age-3 ($N_3$), and an age-4-plus group ($N_4$), with Ricker recruitment on
adult biomass

$$\log R_t = \alpha_0 + \log B_{t-\ell} - b\,B_{t-\ell} - \tfrac{1}{2}\sigma_R^2 + \varepsilon_t$$

where $B_t = N_{3,t}w_3 + N_{4,t}w_4$ is standing adult biomass in grams and
$\ell = `r Vi(CFG$biology$recruit_lag)`$ years, and density-dependent apparent
survival

$$\mathrm{logit}(s_t) = \phi_0 + \phi_N\,N_{\mathrm{all},t-1}/100 + \zeta_t .$$

**The posterior used here contains no flow covariate.** This is the circularity
guard: were abundance estimated under a flow-informed model, a flow signal would be
embedded in the response before any regression was fitted.

Two responses were constructed, each on all
`r Vi(RESP$meta$n_draws[1])` retained posterior draws so that abundance uncertainty
propagates rather than being collapsed to a point estimate:

**Recruitment.** $y_t^{(m)} = \log R_t^{(m)} - \log B_{t-\ell}^{(m)}$. Stock is
adult **biomass**, matching the IPM's own Ricker.

**Survival.** The process deviate
$\varepsilon_t^{(m)} = \mathrm{logit}(s_t^{(m)}) - \phi_0^{(m)} - \phi_N^{(m)} N_{\mathrm{all},t-1}^{(m)}/100$,
i.e. the part of survival the density-dependent null model leaves unexplained.
Regressing survival itself on discharge would re-attribute density-driven variation
to flow, because flow drives recruitment and recruitment drives density at a lag.

## Hydrograph standardization

Daily discharge was log-transformed, centred on a day-of-year climatology, and
scaled by a single global standard deviation:

$$x_y(t) = \frac{\log Q_y(t) - \bar{c}(t)}{\sigma},\qquad \bar{c}(t) = \frac{1}{n_y}\sum_y \log Q_y(t).$$

The scale is global rather than day-specific; a per-day scale would inflate
low-variance winter days relative to the snowmelt peak. Curves are placed on a
**calendar-year** axis (1 January – 31 December). This axis was chosen for
interpretability: results are reported to water managers, irrigators and
lease-holders who work in calendar years, and a day index anchored to 1 October
is a persistent source of confusion outside the hydrologic community.

The trade-off is stated plainly. Brown trout spawn in autumn and emerge the
following spring, so the spawning-to-emergence interval straddles the calendar
boundary and no single calendar year contains all of it. The analysis accommodates
this through the site-specific flow lag rather than through the axis: a lag equal
to the recruit lag places the hydrograph in the spawning year, while a shorter lag
places it in the rearing year, and the integrated population model selected between
these hypotheses independently at each site. A consequence is that a mechanism
operating continuously across the turn of the year — incubation from December
through March — is split between two lag years and would be detected only weakly in
either. Years without a complete 365-day record were excluded.

## Functional regression

For each posterior draw the response was regressed on the full daily hydrograph,

$$y_t^{(m)} = \mu^{(m)} + \gamma^{(m)} S_{t-\ell} + \sum_{d=1}^{365}\beta^{(m)}(d)\,x_{t}(d) + e_t^{(m)},$$

```{r estimator-methods, results='asis'}
# The estimator description MUST branch. A Methods section that hardcodes
# "P-spline" while the analysis ran fpc describes a model nobody fitted -- and it
# would read as correct to anyone who did not check config.R.
est <- S[["estimator"]] %||% CFG$fitting$estimator
site1 <- CFG$sites$active[[1]]

if (identical(est, "penalized")) {
  cat(sprintf(paste0(
    "with $\\beta(\\cdot)$ a P-spline of basis dimension %s and a second-difference ",
    "roughness penalty, evaluated by unit-weight daily quadrature (the ",
    "summation-convention implementation of `mgcv`; Wood 2017 §7.10). The smoothing ",
    "parameter was estimated once by REML on the posterior-mean response and held ",
    "fixed across draws, so the estimand does not drift between draws. The ",
    "resulting effective degrees of freedom are reported as a **result**: %s for ",
    "recruitment and %s for survival.\n"),
    Vi(CFG$fitting$k_beta),
    V(S[[paste0("edf_recruitment_", site1)]], 1),
    V(S[[paste0("edf_survival_", site1)]], 1)))

} else {
  rule <- S[["fpc_rule"]] %||% CFG$fitting$fpc_rule
  rule_txt <- switch(rule,
    cumulative = sprintf("components were retained until they jointly explained %s%% of the variance in the standardized hydrographs",
                         Vi(S[["fpc_target_var"]] %||% CFG$fitting$fpc_target_var)),
    individual = sprintf("components were retained only if each individually explained at least %s%% of the variance in the standardized hydrographs",
                         Vi(S[["fpc_min_var"]] %||% CFG$fitting$fpc_min_var)),
    both       = sprintf("components were retained until they jointly explained %s%% of hydrograph variance, and only those individually explaining at least %s%%",
                         Vi(S[["fpc_target_var"]] %||% CFG$fitting$fpc_target_var),
                         Vi(S[["fpc_min_var"]] %||% CFG$fitting$fpc_min_var)))

  cat(sprintf(paste0(
    "with $\\beta(\\cdot)$ expanded in the leading functional principal components ",
    "of the standardized hydrographs. Functional PCA was computed on the training ",
    "curves, %s, subject to a maximum of %s. The response was regressed on the ",
    "retained component scores by ordinary least squares and the coefficients ",
    "mapped back to the daily domain as $\\beta(d) = \\sum_k b_k \\phi_k(d)$. ",
    "At %s this retained **%s components**, spanning %s%% of hydrograph variance.\n"),
    rule_txt, Vi(S[["fpc_max"]] %||% CFG$fitting$fpc_max), site1,
    Vi(S[[paste0("fpc_K_", site1)]]), V(S[[paste0("fpc_var_explained_", site1)]], 1)))

  if (isTRUE(S[[paste0("fpc_K_capped_", site1)]] == 1))
    cat("\nThe component cap, rather than the variance rule, determined $K$ at this site; ",
        "the reported $K$ should therefore be read as a ceiling imposed by the analyst ",
        "rather than a property of the data.\n")
  if (isTRUE(S[[paste0("fpc_K_unstable_", site1)]] == 1))
    cat("\nAt least one component fell within 0.5 percentage points of the retention ",
        "threshold, so $K$ varies between cross-validation folds; the validation ",
        "statistics below incorporate that variation because selection was repeated ",
        "within each fold.\n")

  cat(paste0("\nComponent selection is unsupervised with respect to the response: ",
             "components are ordered by variance in the predictor and only then tested ",
             "against recruitment. A mode of flow variation carrying little variance ",
             "but a strong biological signal would be discarded before it was tested ",
             "(Jolliffe 1982). We therefore report the incremental predictive ",
             "contribution of the discarded components in SI, and note that the ",
             "covariance of $\\hat\\beta$ is of rank $K$ and so excludes uncertainty in ",
             "the truncation itself.\n"))
}
```

The hydrograph paired with each recruit year is **not assumed to be the spawning
year**. The integrated population model's indicator-variable selection chose a
flow lag independently at each site, and that choice is carried through here: a
lag equal to the recruit lag identifies spawning-year conditions, while a shorter
lag identifies age-0 rearing-year conditions. On a calendar axis these are
distinct years and the mapping is direct: the hydrograph for recruit year $t$ is
that of calendar year $t - L$, where $L$ is the selected lag. The selected lag is
reported per site rather than assumed (`derived_stats.csv`, `flow_lag_*`).

The **full 365 × 365 covariance** of $\hat\beta$ was retained. Adjacent days are
strongly correlated, so every statement about *where* the curve is high — bands,
contrasts, the location of extrema — requires the off-diagonal terms.

Survival lags of `r paste(CFG$fitting$survival_lags, collapse = " and ")` years were
fitted separately, because adult survival responds to contemporaneous or
prior-year conditions rather than to conditions three years earlier; the selected
lag was `r Vi(S[[paste0("survival_lag_", CFG$sites$active[[1]])]])` and its
permutation $p$-value is Bonferroni-adjusted for the number of lags examined.

## A pre-registered confirmatory model

Because a 365-point curve estimated from
`r Vi(S[[paste0("n_years_", CFG$sites$active[[1]])]])` annual observations spends
degrees of freedom on structure the data may not support, the same responses were
also regressed on the mean standardized anomaly within four pre-registered
calendar-year windows — autumn spawning, overwinter incubation, spring snowmelt,
summer base flow (`r paste(sapply(CFG$windows, function(w) paste(w, collapse = "–")), collapse = "; ")`
on the day of year). This four-parameter model is the powered
confirmatory test of whether flow matters at all; $\beta(d)$ then describes when.
Both were scored on identical folds against identical nulls.

## Validation

Predictive skill was assessed by **blocked** cross-validation holding out
contiguous `r Vi(CFG$fitting$block_len)`-year runs. Leave-one-year-out validation
is inappropriate here: interannual flow is autocorrelated, so a held-out
hydrograph is closely approximated by retained neighbours and the fold measures
interpolation rather than prediction. Both are reported so the magnitude of that
inflation is visible.

Because the null sampling distribution of cross-validated $R^2$ at this sample
size is neither centred at zero nor symmetric, skill was referred to a permutation
null formed by circularly shifting the response against the hydrograph series,
preserving the autocorrelation of both series and breaking only their alignment.
Random shuffling would destroy the autocorrelation and return anticonservative
$p$-values. With
`r Vi(S[[paste0("n_years_", CFG$sites$active[[1]])]])` years the finest resolvable
$p$-value is approximately
`r V(1/(as.numeric(S[[paste0("n_years_", CFG$sites$active[[1]])]]) %||% NA), 3)`.

```{r gate-verdict, results='asis'}
`%||%` <- function(a,b) if (is.null(a)) b else a
if (is.null(GATE)) cat("*[PENDING — run `05_validate.R`.]*\n") else {
  ok_b <- GATE_OK("recruitment","beta"); ok_s <- GATE_OK("recruitment","seasonal")
  if (isTRUE(ok_b) && isTRUE(ok_s))
    cat("Both specifications cleared the gate, and daily-resolution results are reported.\n")
  else if (isTRUE(ok_s))
    cat("The confirmatory four-window model cleared the gate while the daily curve did not. We therefore report that seasonal streamflow affects recruitment but that **its daily timing is not resolvable in this record**, and give guidance at seasonal resolution.\n")
  else
    cat("Neither specification improved on the series mean out of sample. We report the null: seasonal hydrograph structure has no detectable effect on the vital rates at this site over the available record. Curves are presented for completeness and are labelled unsupported.\n")
}
```

## From β(d) to management guidance

β(d) is the effect of a one-standard-deviation standardized log anomaly. Water is
released in cfs. Adding $\Delta Q$ on day $d$ changes the flow signal by

$$\Delta\eta(d) = \frac{\beta(d)}{\sigma}\log\!\left(1 + \frac{\Delta Q}{\bar{Q}(d)}\right),$$

where $\bar{Q}(d)$ is the geometric-mean baseline discharge on that day, so the
benefit per unit of water is approximately $\beta(d)/[\sigma\bar{Q}(d)]$. Because
baseline discharge spans more than an order of magnitude across the calendar year, this
**reshapes** the curve: identical water is a large proportional change at base flow
and a negligible one at the snowmelt peak. We therefore distinguish throughout
between β(d), the ecological sensitivity, and $s(d)$, the management sensitivity.
The logarithm also encodes diminishing returns, so the model implies that
distributing a fixed volume over a longer window outperforms a short pulse.

For a leased volume $V$ released evenly over a window $W$, total benefit is
$\sigma^{-1}\sum_{d\in W}\beta(d)\log(1 + \kappa V/[|W|\bar Q(d)])$, maximised over
window start and length.

The unconstrained optimum is of limited operational use, because water cannot be
moved at arbitrary times of year. We therefore repeated the optimisation
restricted to the season in which allocation decisions are actually made
(`r if (!is.null(DW)) DW$window$label else "the irrigation season"`,
`r if (!is.null(DW)) paste(doy_date(DW$window$start), "to", doy_date(DW$window$end)) else "May to October"`).
This is a constrained optimisation rather than a restricted plot: the best
feasible day is an argmax over a subset and may differ from the global optimum;
its credible interval is recomputed within the subset rather than truncated,
because a simulation whose global peak falls outside the season relocates to its
best feasible day rather than dropping out; and only release windows lying wholly
inside the season are admitted.

We report the resulting **cost of the constraint** — the benefit at the best
feasible day as a fraction of the benefit at the unconstrained optimum. A low
value indicates that the biologically most valuable timing lies outside any window
the allocation system can reach, which is a statement about institutional timing
rather than about water volume.

## Resolution

The location of a maximum was **not** read from the fitted curve. Whole curves were
simulated `r Vi(SE$settings$n_sim)` times from the combined
distribution — a posterior draw for abundance uncertainty plus a multivariate
normal deviate using the full covariance for regression uncertainty — and the
day maximising $s(d)$ recorded for each. The
`r Vi(100 * (SE$settings$cred_level %||% 0.8))`% interval of that distribution is
reported as the resolvable window. Its width is a property of the record rather
than a modelling choice. Simultaneous confidence bands were obtained from the same
simulations as the 95th percentile of the maximum standardized deviation, giving a
critical value of `r V(SE$simultaneous$c_simultaneous[1], 2)` rather than 1.96
(Ruppert, Wand & Carroll 2003 §6.5).

## Total production

Production is a consequence of recruitment and survival rather than a directly
observed response, so it was not regressed on flow. Instead, flow perturbations
were propagated through the age-structured model. Total biomass follows the exact
identity $B_t = B_{t-1} + I_t + G_t - M_t$, partitioning annual change into
recruitment input, somatic growth, and mortality loss, with the age-4-plus group
valued at its abundance-weighted mean mass. There is no harvest term in this model,
so surplus production equals the change in biomass. The identity residual is
computed for every draw and year and is zero to machine precision by construction.

Sensitivity was computed as a paired numerical derivative: for each
`r Vi(PR$settings$block_days)`-day block the projection was run with and without a
`r Vi(PR$settings$delta_cfs)` cfs release under **common random numbers**, so the
difference reflects the release rather than Monte Carlo noise. Runs with each
pathway enabled separately, and both together, isolate the interaction — the
compounding of a survival gain into future spawning stock, which summing the two
curves would miss. Mean mass at age is held at its launch value, so results capture
the flow-on-abundance pathway and **not** temperature-mediated changes in growth.

## Scenario projections

Bureau of Reclamation downscaled daily discharge was standardized against the
**training** climatology and scale, so projected and observed hydrographs share an
axis; standardizing each scenario against itself would re-centre a uniformly drier
future to appear average and eliminate the effect. Scenario effects are
$\eta = \sum_d \beta(d)[x_{\text{scen}}(d) - \bar{x}_{\text{hist}}(d)]$, and are
additionally decomposed by day to identify *which* days drive a scenario's effect —
a scenario may act because flow changes on days the population is sensitive to, or
because it changes greatly on days it is not. The fraction of scenario days falling
outside the training range is reported per scenario.

*[VERIFY: the calendar-year alignment. Recruit year $t$ is paired with calendar year
$t-\ell+1$, on the reasoning that October–November spawning begins a calendar year. A
one-year error shifts β(d) by a full year and remains invisible in the output.]*

## Equilibrium carrying capacity under sustained flow change

The preceding analyses treat flow as a perturbation and ask when it is most
valuable. A separate question, and the one most often asked of reservoir
operations, is where the population settles if summer flow shifts and stays
shifted. We addressed this with a rule-curve analysis, run as an independent arm.

Recruitment and survival were evaluated on a stable-age mapping from spawning
stock to standing abundance,
$X = S/[s(1-s)]$ and $\mathrm{SSB} = S\,w(s)$, giving a closed-form replacement
curve

$$Y(X) = \frac{\exp\!\left(\alpha_0 + \log \mathrm{SSB} - b\,\mathrm{SSB} + \beta_1 z + \beta_2 z^2\right)}{1 - \mathrm{logit}^{-1}\!\left(\phi_0 + \phi_1 z + \phi_2 z^2 + \phi_N X/100\right)},$$

where $z$ is the standardized mean summer discharge. Because survival does not
depend on recruitment, $Y$ is available in closed form and needs no simulation.
Carrying capacity $K$ is the descending intersection of $Y(X)$ with the
replacement line $Y = X$, located by linear interpolation; maximum surplus
production is $\max_X\,[Y(X) - X]$, refined parabolically. Both were computed for
every posterior draw on a common deterministic grid, so that a scenario contrast
reflects flow rather than resampling.

Two features of this arm differ from the rest of the analysis and are deliberate.
First, it uses the **flow-informed** integrated population model rather than the
null fit, because it propagates the flow response that model estimated rather than
estimating a new one; the circularity concern that motivates the null fit
elsewhere does not arise. Second, its covariate is the standardized *annual mean*
summer discharge on the scale the integrated model was fitted with, verified to
reproduce that model's own covariate to within $10^{-10}$; the daily
standardization used for $\beta(d)$ would place the recruitment and survival
coefficients off-scale.

We report the fraction of scenario-years falling outside the calibration range of
$z$, and the location of the vertex of each quadratic flow response. A vertex
inside the observed range with downward curvature implies that the fitted response
already peaks within the historical record.

*[VERIFY before submission: the stable-age weighting $w(s)$ has two conventions in
the source implementation — standing spawning biomass, $w = w_3 + [s/(1-s)]w_4$,
and the alternative that weights deaths, $w = (1-s)w_3 + s\,w_4$. The analysis
selects between them by which reproduces the posterior's own biomass series, not
by preference. State which was used and the agreement obtained.]*

## Data and code availability

All code, standardization constants, and the scripts regenerating every figure and
quoted value are archived at *[REPOSITORY DOI]*. Daily discharge is available from
the USGS National Water Information System (gage 06041000).

---

# SI Appendix

**S1. Circularity guard.** Why the flow-naive posterior is required.

**S2. Estimator comparison.** Daily β(d) against the four-window model on identical
folds and nulls, with leave-one-out values shown alongside and labelled as
interpolation skill.

**S3. Resolution.** Full argmax distributions; the effective-degrees-of-freedom
heuristic against the simulation-based window width.

**S4. Contrasts.** Pairwise $P[s(d_1) > s(d_2)]$ using the full covariance, with the
independence-assuming version for comparison.

**S5. Survival lag sweep.** All lags with unadjusted and adjusted $p$-values.

**S6. Production budget verification.** Maximum relative identity residual across
all draws, years, and scenarios.

**S6b. Allocation-season constraint.** Best feasible release day and its interval
recomputed within the season, against the unconstrained optimum; the fraction of
achievable benefit retained; and the flatness of the benefit surface across
feasible start dates, which determines whether timing is worth negotiating over.

**S6c. Rule-curve arm.** Verification that the summer-mean covariate reproduces the
integrated model's own covariate; agreement between the stable-age geometry and the
posterior biomass series under both weighting conventions; $K$ and $MS$ by
scenario-year with the fraction of draws in which the replacement curve fails to
cross the 1:1 line; quadratic vertex locations relative to the calibration range.

**S7. Sensitivity analyses.** Estimator (penalized versus functional principal
components) and, for the latter, the component-selection rule and threshold
(`fpc_diagnostics_rules.csv`); basis dimension; smoothing parameter fixed versus
per-draw; **the calendar-year domain against a water-year domain**, which bears on
whether an incubation signal spanning the year boundary is being split; stock
covariate retained versus dropped; release size; and for the rule curve, clamped
versus unclamped scenario $z$.

## SI Figures

Standardization (Fig. 2a); baseline discharge and per-cfs leverage (Fig. 2b);
between-year flow variation by day (Fig. 2d); variance decomposition of β(d)
(Fig. 3c); permutation nulls (Fig. 5b); resolution diagnostics (Fig. 5c);
simultaneous versus pointwise bands (Fig. 6a–b); pathway interaction (Fig. 7c);
scenario day attribution (Fig. 8b–c).

## References

Marra & Wood (2012) *Scand. J. Stat.* 39:53–74 · Reiss et al. (2017) *Int. Stat.
Rev.* 85:228–249 · Roberts et al. (2017) *Ecography* 40:913–929 · Ruppert, Wand &
Carroll (2003) *Semiparametric Regression* · Wood (2017) *Generalized Additive
Models*, 2nd ed.
