---
title: "Among-site synthesis"
subtitle: "What daily-scale flow effects reveal about brown trout population ecology in Montana"
---

```{r setup, include=FALSE}
source("_common.R")
```

<!--
=============================================================================
THIS PAGE ANSWERS RESEARCH QUESTION 2.

  RQ1 (results_by_site.html) -- where in the year does water buy the most trout?
  RQ2 (this page)            -- what does that tell us about population ecology?
  RQ3 (scenarios section)    -- how does it change under BoR futures?

RQ2 is the part that makes this a population-ecology paper rather than a
consulting report. The logic: if a beta(t) feature is REPRODUCIBLE ACROSS
INDEPENDENT POPULATIONS, it is evidence about a mechanism. If it appears at one
site only, it is either local hydrology or noise, and 43 years cannot tell those
apart.

Fourteen sites is not 14 replicates of one experiment -- they differ in
regulation, geomorphology, and thermal regime. But they are 14 independent
draws on the same question, and consistency across them is the strongest
inferential lever available without a manipulative experiment.
=============================================================================
-->

# The inferential logic

A single site gives ~43 annual observations. That is not enough to resolve a
365-point curve with confidence, and `results_by_site.html` says so explicitly
for each site.

What multiple sites buy is not more precision at any one site — it is the ability
to ask whether a feature **recurs**. Three outcomes, and they mean different
things:

| Pattern across sites | Interpretation |
|---|---|
| Same β(t) feature, same sign, most sites | Evidence for a shared biological mechanism |
| Feature present but shifted in timing | Mechanism is real, cued locally (elevation, thermal regime, run timing) |
| Feature at one or two sites only | Local hydrology or sampling noise. Do not build a mechanism on it. |

This section reports which of those three applies, for each part of the year, for
recruitment and for survival separately.

## The comparison is only valid within a flow-lag group

The reasoning above assumes every site's β(d) describes **the same biological
year**. That assumption fails whenever the integrated population model selected
different flow lags at different sites, and it did: a lag equal to the recruit lag
puts the hydrograph in the *spawning* year, while a shorter lag puts it in the
*age-0 rearing* year.

Day 200 under a spawning-year lag is the summer **before** spawning. Day 200 under
a rearing-year lag is the first summer of a fish's life. Those are different life
stages, so agreement between them is not evidence of a shared mechanism — and
disagreement is not evidence against one.

Sites are therefore grouped by lag below, and agreement is computed **within**
groups only.

```{r lag-groups}
if (is.null(BR) || is.null(BR$fits)) {
  cat("> Run `R/03_fit_beta_recruitment.R`.\n")
} else {
  lg <- data.frame(
    site = names(BR$fits),
    flow_lag = vapply(BR$fits, function(f) as.integer(f$flow_lag %||% NA), integer(1)),
    row.names = NULL)
  lg$hypothesis <- ifelse(is.na(lg$flow_lag), "unknown",
                   ifelse(lg$flow_lag >= CFG$biology$recruit_lag,
                          "spawning-year conditions", "age-0 rearing-year conditions"))
  kable(lg, caption = "Flow lag selected per site by the IPM. Only sites sharing a hypothesis are directly comparable in the figures below.")
}
```

```{r lag-group-warning, results='asis'}
if (!is.null(BR) && !is.null(BR$fits)) {
  lags <- vapply(BR$fits, function(f) as.integer(f$flow_lag %||% NA), integer(1))
  if (length(unique(na.omit(lags))) > 1)
    cat("\n> **More than one flow lag is in use.** The figures below separate sites by
lag. Do not read across the groups: a feature appearing in both is not
corroboration, because the two curves describe different years of the same
fish's life. If the manuscript needs a single across-site statement, it has to
be made within one lag group, and the other group reported separately.\n")
  else
    cat("\n> All sites share one flow lag, so the curves describe the same biological
year and are directly comparable.\n")
}
```

---

# 1. Do sites share a β(t) shape?

```{r shared-shape, fig.height=6}
if (is.null(BETAR)) {
  cat("> **NOT AVAILABLE** — run `R/03_fit_beta_recruitment.R`.\n")
} else if (dplyr::n_distinct(BETAR$site) < 2) {
  cat("> **ONE SITE ONLY.** This page needs at least two sites.\n>\n",
      sprintf("> Currently fitted: **%s**. Add the rest to `CFG$sites$active` in\n",
              paste(unique(BETAR$site), collapse = ", ")),
      "> `config.R` and re-run — the pipeline already loops.\n")
} else {
  lags <- lag_of(BR)
  # Divide by the site's own SD so this compares SHAPE, not effect size, then
  # facet by lag group: a spawning-year curve and a rearing-year curve describe
  # different years and averaging them would be meaningless.
  cur <- BETAR %>% group_by(site) %>%
    mutate(beta_z = beta / sd(beta, na.rm = TRUE)) %>% ungroup() %>%
    mutate(lag_group = lag_group_of(lags[site]))

  ggplot(cur, aes(doy, beta_z, group = site)) +
    geom_hline(yintercept = 0, linetype = 2, colour = PAL[["mute"]]) +
    geom_line(alpha = 0.45, colour = PAL[["blue"]], linewidth = 0.6) +
    stat_summary(aes(group = lag_group), fun = mean, geom = "line",
                 colour = PAL[["ink"]], linewidth = 1.4) +
    facet_wrap(~ lag_group) +
    scale_x_continuous(breaks = MON_B, labels = MON_L) +
    labs(title = "Recruitment: every site's beta(d), scaled to comparable magnitude",
         subtitle = paste0("Thin blue = one site. Thick black = the mean shape WITHIN a lag group.\n",
                           "Compare within a panel only — panels describe different years of the fish's life."),
         x = NULL, y = "beta(d), scaled within site") +
    theme_wh
}
```

```{r agreement, fig.height=4.2}
if (!is.null(BETAR) && dplyr::n_distinct(BETAR$site) >= 2) {
  lags <- lag_of(BR)
  ag <- BETAR %>% group_by(site) %>%
    mutate(beta_z = beta / sd(beta, na.rm = TRUE)) %>% ungroup() %>%
    mutate(lag_group = lag_group_of(lags[site])) %>%
    group_by(lag_group, doy) %>%
    summarise(consistency = max(mean(beta_z > 0), mean(beta_z < 0)),
              n_sites = dplyr::n(), .groups = "drop")

  ggplot(ag, aes(doy, consistency)) +
    geom_hline(yintercept = 0.5, linetype = 3, colour = PAL[["mute"]]) +
    geom_hline(yintercept = 0.8, linetype = 2, colour = PAL[["red"]]) +
    geom_line(colour = PAL[["blue"]], linewidth = 0.9) +
    facet_wrap(~ lag_group) +
    scale_y_continuous(limits = c(0.4, 1)) +
    scale_x_continuous(breaks = MON_B, labels = MON_L) +
    labs(title = "How often do sites agree on the sign of the flow effect?",
         subtitle = paste0("Fraction of sites on the majority side of zero, computed WITHIN a lag group.\n",
                           "Above the red line (0.8) the effect is consistent enough across independent\n",
                           "populations to argue for a mechanism rather than local hydrology."),
         x = NULL, y = "fraction of sites agreeing on sign") +
    theme_wh
}
```

```{r agreement-text, results='asis'}
if (!is.null(BETAR) && dplyr::n_distinct(BETAR$site) >= 2) {
  lags <- lag_of(BR)
  ag <- BETAR %>% group_by(site) %>%
    mutate(beta_z = beta / sd(beta, na.rm = TRUE)) %>% ungroup() %>%
    mutate(lag_group = lag_group_of(lags[site])) %>%
    group_by(lag_group, doy) %>%
    summarise(cons = max(mean(beta_z > 0), mean(beta_z < 0)),
              dir = ifelse(mean(beta_z) > 0, "positive", "negative"), .groups = "drop")

  for (g in unique(ag$lag_group)) {
    a <- ag %>% filter(lag_group == g) %>% arrange(doy)
    runs <- rle(a$cons >= 0.8)
    ends <- cumsum(runs$lengths); starts <- ends - runs$lengths + 1
    keep <- which(runs$values & runs$lengths >= 14)
    cat(sprintf("\n**%s.** ", g))
    if (length(keep)) {
      cat("Periods where at least 80% of sites agree, lasting two weeks or longer:\n\n")
      for (i in keep)
        cat(sprintf("- **%s to %s** (%d days), effect %s\n",
                    doy_date(a$doy[starts[i]]), doy_date(a$doy[ends[i]]),
                    ends[i] - starts[i] + 1, a$dir[starts[i]]))
    } else {
      cat("No period of two weeks or more reaches 80% agreement. That is a
substantive result, not a gap: it means the *timing* of flow sensitivity is not
shared across these populations, which argues against a single basin-wide release
calendar and for site-specific guidance.\n")
    }
  }
}
```

---

# 2. Recruitment and survival: one mechanism or two?

```{r rec-vs-surv, fig.height=5}
if (is.null(BETAS)) {
  cat("> **NOT AVAILABLE** — run `R/04_fit_beta_survival.R`.\n")
} else {
  both <- bind_rows(
    BETAR %>% mutate(process = "recruitment"),
    BETAS %>% mutate(process = "survival")) %>%
    group_by(process, site) %>%
    mutate(beta_z = beta / sd(beta, na.rm = TRUE)) %>% ungroup()

  ggplot(both, aes(doy, beta_z, group = interaction(site, process), colour = process)) +
    geom_hline(yintercept = 0, linetype = 2, colour = PAL[["mute"]]) +
    geom_line(alpha = 0.35, linewidth = 0.5) +
    stat_summary(aes(group = process), fun = mean, geom = "line", linewidth = 1.3) +
    scale_colour_manual(values = c(recruitment = PAL[["blue"]], survival = PAL[["green"]]),
                        name = NULL) +
    scale_x_continuous(breaks = MON_B, labels = MON_L) +
    labs(title = "Recruitment and adult survival respond to flow at different times of year",
         subtitle = paste0("Thin = individual sites, thick = across-site mean. Both scaled within site.\n",
                           "Different peaks mean two distinct pathways and a release calendar must choose\n",
                           "between them; coincident peaks mean water does double duty.\n",
                           "NOTE: the survival lag is fitted separately from the recruitment flow lag,\n",
                           "so these two curves need not describe the same year."),
         x = NULL, y = "beta(d), scaled within site") +
    theme_wh
}
```

```{r rec-vs-surv-text, results='asis'}
if (!is.null(BETAS) && !is.null(BETAR)) {
  mr <- BETAR %>% group_by(site) %>% mutate(z = beta / sd(beta)) %>%
    group_by(doy) %>% summarise(m = mean(z), .groups = "drop") %>% arrange(doy)
  ms <- BETAS %>% group_by(site) %>% mutate(z = beta / sd(beta)) %>%
    group_by(doy) %>% summarise(m = mean(z), .groups = "drop") %>% arrange(doy)
  r <- cor(mr$m, ms$m, use = "complete.obs")
  cat(sprintf(paste0("Across-site mean curves peak on **%s** for recruitment and **%s** for ",
                     "survival, and are correlated at r = **%.2f** across the year.\n\n"),
              doy_date(mr$doy[which.max(mr$m)]), doy_date(ms$doy[which.max(ms$m)]), r))
  if (abs(r) < 0.3)
    cat("The two pathways are close to independent. Water released for recruitment is not
the same water released for adult survival, and they should be presented as separate
management levers with separate calendars.\n")
  else if (r > 0.5)
    cat("The two pathways move together. A single release window serves both — the best
case for management, and worth stating plainly.\n")
  else
    cat("Partial overlap. There is a compromise window, and the production sweep in
`results_by_site.html` reports the joint optimum, which is the right quantity because
it propagates both pathways through the age structure.\n")
}
```

---

# 3. Does the shared structure line up with brown trout life history?

This is the ecological test. β(t) is estimated with no biological input beyond
the lag structure — it does not know when trout spawn. If the days it flags line
up with known life-history transitions, that is independent corroboration. If
they do not, either the mechanism is something other than the obvious one, or the
curve is fitting noise.

```{r life-history, fig.height=5.5}
if (is.null(LIFE) || is.null(STAGE)) {
  cat("> **NOT AVAILABLE** — run `R/09_figures.R`, which writes",
      "`life_history_timeline.csv` and `life_stage_sensitivity.csv`.\n>\n",
      "> **Edit the `LIFE` timeline in `R/09_figures.R` to match your system and cite the source.**",
      "The alignment test is only as good as that timeline.\n")
} else {
  kable(STAGE, digits = 3,
        caption = paste0("Mean flow sensitivity within each life-history stage, averaged across sites. ",
                         "Stage boundaries come from the timeline in R/09_figures.R -- edit and cite them."))
}
```

```{r life-history-caveat, results='asis'}
if (!is.null(STAGE)) {
  cat("\n> **Read this before quoting the table.** The stage boundaries were set by hand, ",
      "not estimated. Choosing them after seeing β(t) would guarantee agreement and prove ",
      "nothing. Fix the timeline from the literature first, then compare — and say in the ",
      "manuscript that you did it in that order.\n")
}
```

---

# 4. Does site-to-site variation track anything?

```{r site-covariates, results='asis'}
cat("If the flow response differs among sites, the next question is whether it differs
*systematically*. Candidate explanations, in rough order of how testable they are
with what you have:\n\n")
cat("| Candidate | Why it could matter | Data needed |\n")
cat("|---|---|---|\n")
cat("| Degree of regulation | A dam-controlled reach has a compressed natural flow range, so fish may be less flow-limited | Reservoir presence, distance downstream |\n")
cat("| Elevation / thermal regime | Sets spawning and emergence timing, which shifts β(t) horizontally | Site elevation, summer temperature |\n")
cat("| Baseline discharge | Determines how much a given release changes proportional flow — see Proposal 02 | Already computed: per-day climatology |\n")
cat("| Channel confinement | Governs whether high flow means habitat gain or scour | Geomorphic classification |\n")
cat("| Irrigation withdrawal | Sets whether summer low flow is naturally or anthropogenically limiting | Diversion records |\n\n")
cat("**None of these is in the pipeline yet, because none is in the data folder.** ",
    "With 14 sites you can support one, possibly two site-level covariates — not five. ",
    "Pick the one with the strongest mechanistic prior *before* fitting, and treat the ",
    "rest as future work. Fourteen sites and five covariates is the same overfitting ",
    "problem as 43 years and a 365-point curve, one level up.\n")
```

```{r discharge-driver, fig.height=4.5}
# Columns come from sensitivity_daily.csv (script 06): doy, Qbar_cfs, effect,
# sens_per_delta. An earlier draft referenced `delta_logRS` and `Q_baseline`,
# which that script has never written.
if (!is.null(DAILY) && dplyr::n_distinct(DAILY$site) >= 2) {
  bd <- DAILY %>% filter(process == "recruitment") %>%
    group_by(site) %>%
    summarise(best_effect = max(effect, na.rm = TRUE),
              median_Q = median(Qbar_cfs, na.rm = TRUE), .groups = "drop")

  ggplot(bd, aes(median_Q, best_effect)) +
    geom_point(size = 3, colour = PAL[["blue"]]) +
    geom_text(aes(label = site), size = 2.8, vjust = -1, colour = "grey30") +
    scale_x_log10() +
    labs(title = "Do smaller streams get more benefit per cfs released?",
         subtitle = paste0("A prediction that falls straight out of the model structure: benefit per cfs\n",
                           "scales as 1/Q, so low-flow systems should show larger effects for the same\n",
                           "water. If this slopes down, that is a mechanism confirmed rather than a\n",
                           "nuisance — and it tells managers where to spend leased water first."),
         x = "median discharge (cfs, log scale)",
         y = paste0("largest per-", unique(DAILY$delta_cfs)[1], "-cfs benefit (%)")) +
    theme_wh
}
```

---

# 5. The confirmatory anchor across sites

```{r pooled-anchor}
if (is.null(GATE) || !"seasonal_r2_blocked" %in% names(GATE)) {
  cat("> **NOT AVAILABLE** — run `R/05_validate.R`.\n")
} else {
  kable(GATE, digits = 3,
        caption = paste0("Validation for every site: the functional beta(t) model and the four-window ",
                         "seasonal anchor, on identical blocked folds with identical permutation nulls."))
}
```

```{r pooled-text, results='asis'}
if (!is.null(GATE)) {
  np <- sum(GATE$passes, na.rm = TRUE); ns <- nrow(GATE)
  cat(sprintf("\n**%d of %d sites clear the gate.**\n\n", np, ns))
  if (np == 0) {
    cat("No site individually shows a flow signal that survives blocked cross-validation and ",
        "a circular-shift permutation null. Before concluding there is nothing here, note two ",
        "things.\n\n",
        "First, **the pooled test is the one with power.** Fourteen underpowered site tests can ",
        "all fail while a hierarchical fit detects a shared effect, because pooling estimates one ",
        "common quantity from all the data at once. Section 1 above is the informal version; the ",
        "formal version is a hierarchical model on the seasonal anchor, which is the next thing ",
        "to build.\n\n",
        "Second, **reporting the subset that passed would be a selection problem** even if some ",
        "had. At α = 0.05 across 14 sites you expect roughly one pass by chance. The pooled ",
        "estimate has no selection step, which is why it is the defensible headline.\n")
  } else if (np < ns) {
    cat("A mixed result is the most interesting outcome and the hardest to write up honestly. ",
        "The temptation is to report the sites that passed. Resist it: with 14 tests at ",
        "α = 0.05 you expect about one pass by chance, so a subset of passes is not evidence ",
        "that those particular sites are special.\n\n",
        "The defensible framing is the pooled effect plus the between-site SD. A large mean ",
        "with large between-site variation is a real finding — *on average yes, but not ",
        "everywhere* — and it is exactly what a basin-scale manager needs to know.\n")
  } else {
    cat("Every site clears the gate independently. That is strong, and it makes the ",
        "across-site consistency in Section 1 the natural headline: the same seasonal ",
        "structure recovered from independent populations.\n")
  }
}
```

---

# 6. What this says about Montana brown trout population ecology

```{r ecology-synthesis, results='asis'}
if (is.null(BR)) {
  cat("> Run the pipeline first.\n")
} else {
  cat("Draft the synthesis around whichever of these the results actually support.
None should be asserted without the corresponding figure above backing it.\n\n")
  cat("1. **Which vital rate is flow-limited, and when.** If recruitment β(t) is
structured and survival β(t) is flat, brown trout population dynamics here are
recruitment-driven and adult survival is buffered against hydrology. That is a
statement about where density dependence and environmental forcing act, and it
has direct consequences for what management can and cannot change.\n\n")
  cat("2. **Whether the limiting window is natural or anthropogenic.** A negative
spring-runoff effect is a natural-disturbance signal (fry displacement, redd
scour). A positive summer base-flow effect in a diverted system is an
anthropogenic-limitation signal. These call for opposite management responses:
the first argues for accommodating natural variability, the second for
augmenting it.\n\n")
  cat("3. **Whether timing is cued locally.** If the β(t) shape is shared but shifted
between sites, the mechanism is real and phenological — and the shift should
correlate with elevation or thermal regime. That is a testable prediction from
this analysis, and a good final paragraph.\n\n")
  cat("4. **The resolution limit is itself a result.** If 43 years cannot resolve
finer than a ~60-day window, that bounds how precisely *any* flow-management
prescription can be specified from observational data at this sample size. Saying
so quantitatively is more useful to the field than a spuriously precise date, and
it motivates the experimental releases that would resolve it.\n")
}
```

---

# 7. Limits

```{r limits, results='asis'}
cat("State these in the manuscript. A reviewer will find them anyway, and finding
them already stated is a different experience from finding them omitted.\n\n")
cat("- **Sites are not independent replicates.** Several share a basin, a climate signal, and in some cases a reservoir. Shared response may reflect shared weather rather than shared biology. A spatial or basin-level random effect is the fix; it is not yet implemented.\n")
cat("- **The response is a model output, not an observation.** log(R/S) comes from the null IPM's posterior. Section 1 of `methods.html` covers the circularity guard; the residual risk is that IPM structural assumptions are shared across sites and could induce shared β(t) structure.\n")
cat("- **β(t) is linear in the flow anomaly.** A dome-shaped response — too little water bad, too much water bad — would be averaged into a shallow monotone slope. Testable by adding a squared term, but only worth spending the degrees of freedom after the linear model passes.\n")
cat("- **The lag is fixed at ", CFG$biology$recruit_lag, " years and equals the spawning lag**, so flow and spawning stock share a calendar year and any pathway from spawn-year flow to adult abundance is differenced out of log(R/S). This estimates a *spawn-year* effect, not an age-0 rearing-year effect.\n", sep = "")
cat("- **Mean mass at age is frozen** in the production sweep, so temperature effects on growth are outside the model. Production sensitivities are the flow-on-abundance pathway only.\n")
cat("- **Fourteen sites supports one site-level covariate, not five.** See Section 4.\n")
```

---

```{r session, results='asis'}
cat("<hr>\n<small>")
cat(sprintf("Rendered %s · %s", format(Sys.time(), "%Y-%m-%d %H:%M"), R.version.string))
gs <- suppressWarnings(tryCatch(system("git rev-parse --short HEAD", intern = TRUE),
                                error = function(e) NA))
if (length(gs) && !is.na(gs[1])) cat(sprintf(" · commit %s", gs[1]))
cat("<br>Every number on this page is read from <code>output/</code>. ")
cat("Sections showing <strong>NOT AVAILABLE</strong> mean that stage has not been run.")
cat("</small>\n")
```
